An Embedding Framework for Consistent Polyhedral Surrogates

### 2021-09-15 poly-embed-journal/poly-embed-journal.tex
Fleshing out the conclusion and resolving some comments

%\btw{Cut summary for now (Why is this cool?)}
%\jessiet{Jessie: added a bit to say why this is cool.}
\jessiet{Is this following paragraph necessary here?  It's more of an intro for \S~\ref{sec:calibration}.}
We will see that embedding implies the existence of a consistent surrogate and link, so this result tells us that, for discrete losses, the embedding framework is sufficient to understand the existence of consistent surrogates for the given loss.
Even if there is a smooth surrogate for a given discrete loss, there is also a polyhedral loss that is calibrated for the given loss.
What's more is that there is a polyhedral \emph{embedding} that is calibrated for the discrete loss.

Recall from the introduction Theorem~\ref{thm:embed-poly-informal}.
\embedpolyinformal*
This result is exactly the chaining of theorems \ref{thm:poly-embeds-discrete} and \ref{thm:discrete-loss-poly-embeddable}, thus we have shown the result.

@@

\raf{Reference Theorem~\ref{thm:link-informal}}

@@

\linkinformal* \jessie{vs...}

\begin{theorem}\label{thm:eps-thick-calibrated}
  Let $L$ be polyhedral and $\ell$ the discrete loss it embeds from Theorem~\ref{thm:poly-embeds-discrete}.
  Then there exists a calibrated link function $\psi$ from $L$ to $\ell$.
%Then for small enough $\epsilon > 0$, the $\epsilon$-thickened link $\psi$ is well-defined and, furthermore, is a calibrated link from $L$ to $\ell$.
\end{theorem}

@@

While we have focused on polyhedral losses thus far, it turns out that many of our results about embeddings extend to losses with polyhedral Bayes risks.
(We say a concave function is polyhedral if its negation is a polyhedral convex function.)
To see that this is a weaker condition, recall that Theorem~\ref{thm:poly-embeds-discrete} constructs a finite representative set $\Sc$ for any polyhedral loss $L$, and thus $\risk{L} = \risk{L|_\Sc}$ by Lemma~\ref{lem:loss-restrict}, which is polyhedral.
The condition is strictly weaker, in that the Bayes risk may be polyhedral even if the loss itself is not.
For example, a modified hinge loss $L(r)_y = \max(r^2-1,1-ry)$, \jessiet{Would a figure be helpful?}\raft{Yeah!} which matches hinge loss on the interval $[-1,1]$ but is smooth outside the interval $[-2,2]$, still embeds twice 0-1 loss.

\begin{figure}
	\begin{minipage}{0.47\linewidth}
		\centering
		\includegraphics[width=0.95\linewidth]{figs/modhinge.pdf}
		%		\caption{Bayes risk of the 0-1 loss.}
		%		\label{fig:0-1-br}
	\end{minipage}
	\hfill
	\begin{minipage}{0.47\linewidth}
		\centering		\includegraphics[width=0.95\linewidth]{figs/modhinge-br.pdf}
		%		\caption{Bayes risk of hinge loss.}
		%		\label{fig:hinge-br}
	\end{minipage}
	\caption{(L) Modified (expected) hinge loss for fixed distribution, and (R) Bayes risk of modified hinge still matches the Bayes risk of hinge}
	\label{fig:bayes-risks-01}
\end{figure}

@@

\subsection{A property elicitation perspective}

@@

\raf{Future work and conjectures}
\begin{itemize}
\item \raf{This is false I believe, from BEP surrogate $n=4$ and large (small?) $\alpha$ -- you can add a little curvature between the spokes and it still embeds.  Maybe we can rephrase / reformulate.}
  In fact, we conjecture that \emph{any} loss embedding a discrete $\ell$ must be polyhedral on the convex hull of the embedded reports. 
(The convex hull of the embedded reports follows since any point not in the convex hull will never minimize the expected loss.)

\item 
  Since discrete losses have a finite set of reports, and in turn, minimizers, any surrogate embedding the discrete loss must also have a finite set of unique minimizers.
  This is in turn related to another conjecture about the ``convex envelope'' of embeddings: if $L$ embeds $\ell$ by the embedding $\varphi$, the (polyhedral) surrogate $L'$ such that $L'_y$ is the convex envelope of $\{(\varphi(r),L(r)_y)\}_{r\in\R}$ also embeds $\ell$.

\item
  Relation to geometry of losses / superprediction sets?  Say something about trim being finite...

\item
  For which other classes of surrogates / settings is indirect elicitation equivalent to consistency?

\item
  Characterizing embedding dimension \cite{finocchiaro2020embedding}; does poly dim == embedding dim?  ccdim == embedding dim?

@@

\subsection*{Acknowledgements}
We thank Arpit Agarwal and Peter Bartlett for many early discussions, which led to several important insights.
% to a proof of %\raf{1-d reduction} among other insights.
% Additionally, we thank Arpit Agarwal for early insights and bringing out attention to the abstain property.
We thank Eric Balkanski for help with a lemma about submodular functions. %Lemma~\ref{lem:bar-f}\jessie{We don't have the Lov\'asz hinge appendix here?}.
This material is based upon work supported by the National Science Foundation under Grants No.\ 1657598 and No.\ DGE 1650115.
\raf{Nishant, anyone else now?}\jessie{Stephen}
\newpage
\bibliographystyle{plainnat}
\bibliography{diss,extra}

@@

\section{Top-$k$ surrogate} 
\jessie{Remove this section?}
Consider the surrogate and discrete loss $L^k(u)_y~=~\left(\frac{1}{k} \sum_{i=1}^k (u + \ones - e_y)_{[i]} - u_y \right)_+$ given in Equation~\ref{eq:L-2-surrogate}.

\begin{lemma}\label{lem:top-k-polyhedral}
$L^k$ is a polyhedral loss.
\end{lemma}
\begin{proof}
Observe $L^k$ can be written as the pointwise max of $\binom{n}{k} +1$ terms, where the $\binom{n}{k}$ terms are selecting the $k$ elements of $u + \ones - e_y$, and the max comes from selecting the $u_i$ elements with highest weight.
\end{proof}
%By Lemma~\ref{lem:poly-loss-poly-risk}, we then know that $\risk{L'}$ is also polyhedral.

%We next show that $\U = \{u \in \{0,1\}^n : \|u\|_0 \leq k\}$ is always represented among the optimizers of $L'$. \begin{lemma}\label{lem:top-k-optimal-corners}
%  For all $p \in \simplex$ we have $\prop{L'}(p) \cap \U \neq \emptyset$.
%\end{lemma}
%\begin{proof}
%  Let us first simplify $L'$:
%  \[
%    L'(u)_y = \left( 1 - u_y + \frac{1}{k} \sum_{i=1}^k (u - e_y)_{[i]} \right)_+~.
%  \]
%  Examining this expression, observe that
%  if $u_i < u_{[k+1]}$, then $(u-e_y)_i < (u-e_y)_{[k]}$, as only one entry changes (and decreases) from one to the other.
%  Hence, if $u_i < u_{[k+1]}$, then the index $i$ will not appear in the summation for any $y$, and thus only influences the loss when $y=i$.
%  Now consider the loss of $u'$ which is the same as $u$ but with $u_i' = u_{[k+1]}$.
%  Then $L'(u')_y \leq L'(u)_y$ when $y=i$ and $L'(u')_y = L'(u)_y$ when $y\neq i$.
%  From this we see that without loss of generality, $u_{[k+1]} = \cdots = u_{[n]}$.
%
%  Next, we show that $L'$ is invariant in the $\ones$ direction:
%  \begin{align*}
%    L'(u + \alpha \ones)_y &= \left( 1 - (u_y + \alpha) + \frac{1}{k} \sum_{i=1}^k ((u + \alpha) - e_y)_{[i]} \right)_+\\
%    &= \left( 1 - (u_y + \alpha) + \frac{1}{k}\left[ \alpha k + \sum_{i=1}^k (u - e_y)_{[i]} \right] \right)_+\\
%    &= \left( 1 -u_y + \frac{1}{k}\left[ \sum_{i=1}^k (u - e_y)_{[i]} \right] \right)_+\\
%  &= L'(u)_y~.
%  \end{align*}
%  Thus, without loss of generality, we now have $u_{[k+1]} = \cdots = u_{[n]} = 0$.
%
%  Let $u_{[1..j]} := \sum_{i=1}^j u_{[i]}$.
%  We now claim that for all $y \in \Y$, the minimum of $L(u)_y$ is achieved by some $u \in [0,1]^n$.
%  \raf{Only hole left is the following line.  I think it might not be true.}
%  This follows from the observation that if $u_i > 1$, the the loss only decreases by reporting $u'$ which is equal to $u$ but with $u'_i = 1$.
%  For $u\in[0,1]^n$, the argument to $(\cdot)_+$ is nonnegative, so we can rewrite the expected loss of $L'$ as
%  \begin{align*}
%  	\inprod{p}{L'(u)} &= 1 - \inprod{p}{u} + \frac{1}{k} \left( \sum_{i=1}^k (1 - p_{[i]}) u_{[i]} \right) = 1 + \inprod{\tfrac 1 k \ones - (1+\tfrac 1 k)p}{u}~.
%  \end{align*}
%  As a linear objective over the hypercube, $\inprod{p}{L'}$ must be optimal on some corner of the hypercube, so we conclude $u_{[1]}, u_{[2]}, \ldots, u_{[k]} \in \{0,1\}$.
%  In particular, for all $p \in \simplex$, there is a report $u \in \{0,1\}^n$ optimizing $L'$; since we can take $\|u\|_0 \leq k$ by the above observation that $u_{[k+1]} = u_{[k+2]} = \ldots = u_{[n]} = 0$, some $u \in \U$ must be optimal as well.
%\end{proof}
%
%\raf{Cut this final lemma for now}
% \begin{lemma}\label{lem:top-k-surrogate-embeds}
% $L':\reals^n \to \reals^\Y_+$ embeds $\ell':\R \to \reals^\Y_+$.
% \end{lemma}
% \begin{proof}
%   We can verify $\ell'(r) = L'(\varphi(r))$ for all $r \subseteq [n]$ and $\varphi(r)$ being its conversion from a set to binary vector, so $\varphi(\R) = \{ u \in \{0,1\}^n : \|u\|_1 \leq k \}$.
%   Moreover, by Lemma~\ref{lem:top-k-optimal-corners} and the above equality, respectively, we know that $\inf_{u \in \reals^n} \inprod{p}{L'(u)} = \min_{u \in \varphi(\R)} \inprod{p}{L'(u)} = \min_{r \in \R}\inprod{p}{\ell'(r)}$ for all $p \in \simplex$, and so we conclude that $L'$ embeds $\ell'$.
% \end{proof}

@@

%\raf{Let's use $\H^*$ for the minimum one, here and below; $\H$ for another one}
%\raf{Similarly $\V^*$ for the minimum one, below; $\V$ for another one}
\begin{theorem}\label{thm:polyhedron-uniquely-gen-facets}
  Given a $d$-dimensional polyhedron $P \subseteq \reals^d$, there is a unique finite set of closed halfspaces $\H^*$ such that $P = \cap \H^*$, and for all finite sets of closed halfspaces $\H$ such that $P = \cap \H$, we have $\H^* \subseteq \H$.
  Moreover, $\{H\cap P \mid H^+ \in \H^*\}$ is the set of facets of $P$.
\end{theorem}
\begin{proof}
	Since $P$ is $d$-dimensional in $\reals^d$, it therefore has nonempty interior.	
  \jessiet{\url{https://arxiv.org/abs/0805.0292} Prop 4.5?}
\end{proof}

>>>>

%\btw{Cut summary for now (Why is this cool?)}
%\jessiet{Jessie: added a bit to say why this is cool.}
\jessiet{Is this following paragraph necessary here?  It's more of an intro for \S~\ref{sec:calibration}.}
We will see that embedding implies the existence of a consistent surrogate and link, so this result tells us that, for discrete losses, the embedding framework is sufficient to understand the existence of consistent surrogates for the given loss.
Even if there is a smooth surrogate for a given discrete loss, there is also a polyhedral loss that is calibrated for the given loss.
What's more is that there is a polyhedral \emph{embedding} that is calibrated for the discrete loss.
Observe that the concatenation of theorems \ref{thm:poly-embeds-discrete} and \ref{thm:discrete-loss-poly-embeddable} yield Theorem~\ref{thm:embed-poly-informal}, and highlight the connection between embeddings and polyhedral losses.

@@

@@

\linkinformal* 
%\begin{theorem}\label{thm:eps-thick-calibrated}
%  Let $L$ be polyhedral and $\ell$ the discrete loss it embeds from Theorem~\ref{thm:poly-embeds-discrete}.
%  Then there exists a calibrated link function $\psi$ from $L$ to $\ell$.
%%Then for small enough $\epsilon > 0$, the $\epsilon$-thickened link $\psi$ is well-defined and, furthermore, is a calibrated link from $L$ to $\ell$.
%\end{theorem}

@@

While we have focused on polyhedral losses thus far, it turns out that many of our results about embeddings extend to losses with polyhedral Bayes risks.
(We say a concave function is polyhedral if its negation is a polyhedral convex function.)
To see that this is a weaker condition, recall that Theorem~\ref{thm:poly-embeds-discrete} constructs a finite representative set $\Sc$ for any polyhedral loss $L$, and thus $\risk{L} = \risk{L|_\Sc}$ by Lemma~\ref{lem:loss-restrict}, which is polyhedral.
The condition is strictly weaker, in that the Bayes risk may be polyhedral even if the loss itself is not.
For example, a modified hinge loss $L(r)_y = \max(r^2-1,1-ry)$ as shown in Figure~\ref{fig:modified-hinge}, which matches hinge loss on the interval $[-1,1]$ but is smooth outside the interval $[-2,2]$, still embeds twice 0-1 loss.

\begin{figure}
	\begin{minipage}{0.47\linewidth}
		\centering
		\includegraphics[width=0.95\linewidth]{figs/modhinge.pdf}
		%		\caption{Bayes risk of the 0-1 loss.}
		%		\label{fig:0-1-br}
	\end{minipage}
	\hfill
	\begin{minipage}{0.47\linewidth}
		\centering		\includegraphics[width=0.95\linewidth]{figs/modhinge-br.pdf}
		%		\caption{Bayes risk of hinge loss.}
		%		\label{fig:hinge-br}
	\end{minipage}
	\caption{(L) Modified (expected) hinge loss for fixed distribution, and (R) Bayes risk of modified hinge still matches the Bayes risk of hinge}
	\label{fig:modified-hinge}
\end{figure}

@@

\subsection{A property elicitation perspective on trimmed losses}
\jessie{Should this be a subsubsection?}

@@

\jessie{New version...?}
In this paper, we introduce the embeddings tool for studying finite prediction problems, and connect this tool to the design of piecewise-linear and convex (polyhedral) surrogates (Theorem~\ref{thm:embed-poly-informal}), and show that embeddings by polyhedral surrgates are consistent with respect to the given target problem (Theorem~\ref{thm:link-informal}).
We additionally show that embedding discrete prediction tasks is tightly connected to surrogates with polyhedral Bayes risks (Corollary~\ref{cor:poly-risk-fin-rep}).
In \S~\ref{sec:applications}, this tool is applied to a variety examples, including high-confidence classification, ordered partitions, top-$k$ prediction, and structured prediction.
Finally, when restricting to polyhedral surrogate losses, we show that indirect property elicitation is equivalent to consistency in \S~\ref{sec:poly-ie-consistency}.

Some open questions remain:

\jessie{I don't think we want all of these... I feel like it sends a signal about the work being incomplete.}
\raf{Future work and conjectures}
\begin{itemize}
\setlength \itemsep{0.1em}
\item %\raf{This is false I believe, from BEP surrogate $n=4$ and large (small?) $\alpha$ -- you can add a little curvature between the spokes and it still embeds.  Maybe we can rephrase / reformulate.}
%  In fact, we conjecture that \emph{any} loss embedding a discrete $\ell$ must be polyhedral on the convex hull of the embedded reports. 
%(The convex hull of the embedded reports follows since any point not in the convex hull will never minimize the expected loss.)
\btw{Modified from a conjecture Raf thought was false. Old version in comments - J, 15 Sept 21}
While the modified hinge example in Figure~\ref{fig:modified-hinge} shows that not every embedding must necessarily be polyhedral, this curvature is added on a set that is never uniquely optimal.
A natural question is to posit if an embedding my be polyhedral on the convex hull of the embedded reports; we conjecture negatively, though some conditions on linearity seem necessary.

\item 
  Since discrete losses have a finite set of reports, and in turn, minimizers, any surrogate embedding the discrete loss must also have a finite set of unique minimizers.
  This is in turn related to another conjecture about the ``convex envelope'' of embeddings: if $L$ embeds $\ell$ by the embedding $\varphi$, the (polyhedral) surrogate $L'$ such that $L'_y$ is the convex envelope of $\{(\varphi(r),L(r)_y)\}_{r\in\R}$ also embeds $\ell$.
  \jessie{What is the significance of this conjecture?}

\item
  Since polyhedral surrogates have a finite trimmed loss, it is an open line of work to relate embeddings to the to geometry of losses~\citep{williamson2014geometry} through superprediction sets, as these sets encapsulate losses that are dominated by elements of $\trim(L)$.

\item
  In \S~\ref{sec:poly-ie-consistency}, we show that indirect elicitation is equivalent to consistency when restricting to the class of polyhedral surrogates.
  It remains a line of future work to characterize other settings in which indirect elicitation and consistency are equivalent.

\item
  \citet{finocchiaro2020embedding} introduce and present lower bounds for the notion of \emph{embedding dimension} of a target loss $\ell$: the minimum prediction dimension $d$ such that a polyhedral surrogate $L:\reals^d\to \reals^\Y_+$ embeds $\ell$.
  However, it is unclear if convex consistency dimension restricting to polyhedral losses, embedding dimension, or unrestricted convex consistency dimension are equal for all target losses.
  %Characterizing embedding dimension \cite{finocchiaro2020embedding}; does poly dim == embedding dim?  ccdim == embedding dim?

@@

\subsection*{Acknowledgements}
We thank Arpit Agarwal and Peter Bartlett for many early discussions, which led to several important insights.
% to a proof of %\raf{1-d reduction} among other insights.
% Additionally, we thank Arpit Agarwal for early insights and bringing out attention to the abstain property.
We thank Eric Balkanski for help with a lemma about submodular functions, Stephen Becker for his insights about finding a constant rate of separation in proving $\epsilon$-thickenings are calibrated.
 %Lemma~\ref{lem:bar-f}\jessie{We don't have the Lov\'asz hinge appendix here?}.
This material is based upon work supported by the National Science Foundation under Grants No.\ 1657598 and No.\ DGE 1650115.
\raf{Nishant, anyone else now?}
\newpage
\bibliographystyle{plainnat}
\bibliography{diss,extra}

@@

@@

%\raf{Let's use $\H^*$ for the minimum one, here and below; $\H$ for another one}
%\raf{Similarly $\V^*$ for the minimum one, below; $\V$ for another one}
\begin{theorem}\label{thm:polyhedron-uniquely-gen-facets}
  Given a $d$-dimensional polyhedron $P \subseteq \reals^d$, there is a unique finite set of closed halfspaces $\H^*$ such that $P = \cap \H^*$, and for all finite sets of closed halfspaces $\H$ such that $P = \cap \H$, we have $\H^* \subseteq \H$.
  Moreover, $\{H\cap P \mid H^+ \in \H^*\}$ is the set of facets of $P$.
\end{theorem}
\begin{proof}
	Since $P$ is $d$-dimensional in $\reals^d$, it therefore has nonempty interior.	
  \jessiet{\url{https://arxiv.org/abs/0805.0292} Prop 4.5? Ziegler Lectures on Polytopes Theorem 2.15 (7) also kind of close to this, but on polytopes not polyhedron.}
\end{proof}
